---
title: "MAA_2022"
author: "Jay Kim"
date: "1/20/2022"
output: 
  html_document:
   toc: true
   toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r datashaping, include=FALSE}
library(readr)
library(tidyverse)
CSE_ALL_ENR_up2020 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/CSE_ALL_ENR_up2020.csv")
FTIC_IDs <- CSE_ALL_ENR_up2020 %>%  filter(COHORT_YEAR>= 20172018) %>% select("UWFID"=UNIV_ROW_ID)
# Hmisc::describe(CSE_ALL_ENR_up2020) 
# colnames(CSE_ALL_ENR_up2020)
# colSums(is.na(CSE_ALL_ENR_up2020))
# addmargins(table(CSE_ALL_ENR_up2020$COHORT_YEAR, CSE_ALL_ENR_up2020$APR))
# high school code
 
applicants_2017to2021_COM_HSGPA2021V1 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/_DataShaping/applicants_2017to2021_COM_HSGPA2021V1.csv")

FL_HS_CEEB <- applicants_2017to2021_COM_HSGPA2021V1 %>% select("hs_ceeb"=HS_CEEB,"HIGH_SCHOOL_NAME"=HS_NAME) %>% unique()
write.csv(FL_HS_CEEB, "FL_HS_CEEB.csv", row.names = F)

mydata.one <- merge(CSE_ALL_ENR_up2020, FL_HS_CEEB, by="HIGH_SCHOOL_NAME", all.x = T) 
 
#school grades
library(readxl)
SchoolGrades21 <- read_excel("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/2022 Active Projects/01/MAA2022/SchoolGrades21.xlsx", 
    sheet = "SG")
# colnames(SchoolGrades21)
SchoolGrades21 <- janitor::clean_names(SchoolGrades21) %>% mutate(high_school = tolower(school_name)) %>% 
    select(high_school,total_points_earned )
 
SchoolGrades21_CEEB<- read_excel("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/2022 Active Projects/01/MAA2022/SchoolGrades21.xlsx", 
    sheet = "CEEB") 
SchoolGrades21_CEEB <- janitor::clean_names(SchoolGrades21_CEEB) %>% 
    mutate(high_school= tolower(high_school))

 
HS_CEEB_POINT <- merge(SchoolGrades21, SchoolGrades21_CEEB, by="high_school", all.y = T)  #1715
 
FL_HS_CEEB_re <- merge(FL_HS_CEEB, HS_CEEB_POINT, by="hs_ceeb" )
 
CSE_ALL <- merge(CSE_ALL_ENR_up2020, FL_HS_CEEB_re, by="HIGH_SCHOOL_NAME", all.x = T) 

CSE_ALL_MAA2022 <- CSE_ALL %>% 
    select(1,19:81)  
#HSGPA mean
mean(CSE_ALL_MAA2022$GPA_HIGHSCHOOL, na.rm=T) #3.826968
fivenum(CSE_ALL_MAA2022$GPA_HIGHSCHOOL)  #0.00 3.45 3.86 4.23 5.04
CRSCSEData <- read_csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202001/CRSCSEData.csv") %>% 
      mutate(Summer = substr(DEMO_TIME_FRAME, 5,6)) %>% filter(Summer == "05" & DEMO_TIME_FRAME >= 201508) %>% 
    group_by(UNIV_ROW_ID) %>% filter(DEMO_TIME_FRAME == max(DEMO_TIME_FRAME)) %>% 
    filter(!duplicated(UNIV_ROW_ID)) %>% select(1, Summer)

```


```{r include=FALSE}

app_2017_to_2022_cal_tier_allV0 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/2021 Active Projects/08/ACCEPTANCE RATE/app_2017_to_2022_cal_tier_allV0.csv" )
app_test_score <- app_2017_to_2022_cal_tier_allV0 %>% select(UWFID, BEST_TEST_SCORE)

addmargins(table(app_2017_to_2022_cal_tier_allV0$APP_YEAR))
addmargins(table(CSE_ALL_ENR_up2020$COHORT_YEAR))

# demographic factors for ftic 2017 to 2020
FTIC_test1 <- merge(FTIC_IDs, app_2017_to_2022_cal_tier_allV0, by="UWFID", all.x = T) %>% 
    mutate(TRANSFER_HOURS_EARNED = ifelse(is.na(TRANSFER_HOURS_EARNED), 0, TRANSFER_HOURS_EARNED)) %>% 
    select(UWFID, ADMIT_TYPE,"PROGRAM"=PROGRAM_DESC.x,BEST_HSGPA,BEST_TEST_SCORE,TRANSFER_HOURS_EARNED,  ST,  RACE, GENDER, CNTY) %>% 
    na.omit() #3919
colSums(is.na(FTIC_test1))

# grades and gpa





```





## Data Description

```{r eval=FALSE, include=FALSE}
CSE_ALL_MAA2022_df <- merge(CSE_ALL_MAA2022, CRSCSEData, by="UNIV_ROW_ID", all.x = T) %>%
    mutate(is.Summer = ifelse(is.na(Summer), "No", "Yes")) %>% 
    mutate(GPA_HIGHSCHOOL = ifelse(GPA_HIGHSCHOOL < 2.00, NA, GPA_HIGHSCHOOL)) %>% 
    mutate(HSGPA_ind = ifelse(GPA_HIGHSCHOOL >= 3.80 , "Above3.80", "Below3.80")) %>% 
    filter(!is.na(APR))
 
addmargins(table(CSE_ALL_MAA2022_df$APR, AVEHSGPA=CSE_ALL_MAA2022_df$HSGPA_ind))
hist(CSE_ALL_MAA2022_df$GPA_HIGHSCHOOL)
write.csv(CSE_ALL_MAA2022_df, "CSE_ALL_MAA2022_DF_V0.csv", row.names = F)
GPAY1_MAA2022_DF_V0 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/2022 Active Projects/01/MAA2022/CSE_ALL_MAA2022_DF_V0.csv") %>% 
    select(7:63,65:67)
# colnames(GPAY1_MAA2022_DF_V0)
GPAY1_MAA2022_DF_V1 <-  GPAY1_MAA2022_DF_V0 %>% select(1,3,10,11,12,13,15,21,22,28,29,31:40,43,46,48,50,51,53,57,59,60,APR)
colnames(GPAY1_MAA2022_DF_V1)
GPAY1_MAA2022_DF_V2 <- GPAY1_MAA2022_DF_V1 %>% mutate(ACT_SAT = pmax(ACT_PROPORTION, SAT_PROPORTION)) %>%
    relocate(ACT_SAT, .after = GPA_HIGHSCHOOL) %>% 
    select(-ACT_PROPORTION, -SAT_PROPORTION,  -PSE) %>% 
    mutate(is.Engineering = ifelse(str_detect( ENTRY_PROGRAM, "...Engineering"), "Y","N")) %>% 
     mutate(is.Engineering = ifelse(str_detect( ENTRY_PROGRAM, "Engineering..."), "Y",is.Engineering)) %>% 
    relocate(is.Engineering, .after =ENTRY_PROGRAM ) %>% select(-ENTRY_PROGRAM)
colSums(is.na(GPAY1_MAA2022_DF_V2))
write.csv(GPAY1_MAA2022_DF_V2, "GPAY1_MAA2022_DF_V2.csv", row.names = F)
```

```{r eval=FALSE, include=FALSE}
library(readr)
GPAY1_MAA2022_DF_V2 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/2022 Active Projects/01/MAA2022/GPAY1_MAA2022_DF_V2.csv") #2529
# Hmisc::describe(GPAY1_MAA2022_DF_V2)
hist(GPAY1_MAA2022_DF_V2$total_points_earned)
mean(GPAY1_MAA2022_DF_V2$GPA_HIGHSCHOOL,na.rm = T) #3.783154

HSGPA_scatter <- GPAY1_MAA2022_DF_V2 %>% filter(!is.na(GPA_HIGHSCHOOL)) %>% #filter(HOURS_BROUGHT_TO_UNIVERSITY == 0) %>% 
    mutate(Fall1GPA = ifelse(FIRST_FALL_GPA < 2.77, "Below2.77", "Above2.77")) %>%  
    #filter(GPA_HIGHSCHOOL < 3.40) %>% 
    filter(FIRST_FALL_GPA < 3.00) %>% 
    filter(!(FIRST_FALL_GPA == GPA_ENTERING_SECOND_FALL)) %>% filter(FIRST_FALL_GPA != 0) %>% filter(GPA_ENTERING_SECOND_FALL != 0) %>% 
    mutate(log1stTermGPA =log(FIRST_FALL_GPA)) %>♠% mutate(log1stYearGPA =log(GPA_ENTERING_SECOND_FALL))   
    
HSP <- HSGPA_scatter %>% ggplot(aes( x = FIRST_FALL_GPA , y = log1stYearGPA , color =  GPA_HIGHSCHOOL)) +
    geom_point() + stat_smooth( method =  "lm", se= TRUE)   +
    labs(title="Semi-log:First Fall GPA vs First Year GPA") + xlab("First Fall GPA") +ylab("First Year GPA")

HSPlog <- HSGPA_scatter %>% ggplot(aes( x = log1stTermGPA , y = log1stYearGPA , color =  GPA_HIGHSCHOOL)) +
    geom_point() + stat_smooth( method =  "lm", se= TRUE)   +
    labs(title="Log-log:First Fall GPA vs First Year GPA") + xlab("First Fall GPA") +ylab("First Year GPA")
HSP
HSPlog
shapiro.test(HSGPA_scatter$GPA_ENTERING_SECOND_FALL)

summary(slm1 <- lm(log1stYearGPA ~ FIRST_FALL_GPA, data=HSGPA_scatter  ))
summary(slm2 <- lm(log1stYearGPA ~ log1stTermGPA, data=HSGPA_scatter  ))



```

### Better Data with each term GPA

```{r}
lastvalue = function(x) tail(x[!is.na(x),1])
STEM=c("STEM")
heslth=c("Health", "Gap Analysis")
#test_score <- applicants_2017to2021_COM_HSGPA2021V1 %>% select(UWFID, BESTTES)
dffGPA_varis <- c(#"DIFFGPA",    
                  "FallWHours","SpringWHours", "SwitchedMajor","Cohort",  "BEST_TEST","Prior_Hours","AREA_OF_STRATEGIC_EMPHASIS",
                  "AGE_AT_ENTRY",
                  "GENDER","COUNTY_GROUP","GPA_HIGHSCHOOL","FIRST_FALL_PELL","FIRST_FALL_BRIGHT_FUTURES","FIRST_GEN_IND","AGE_AT_ENTRY")
library(readr)
GPA_UWF_MAAV_majorAPR <- read_csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/_DataShaping/GPA_UWF_MAAV_majorAPRONLY.csv")
GPA_UWF_ADD_DIFFGPAV0 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/_DataShaping/GPA_UWF_ADD_DIFFGPAV0.csv") # updated
GPA_UWF_ADD_DIFFGPAV1 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/_DataShaping/GPA_UWF_ADD_DIFFGPAV1.csv") #updated fall hours


GPA.DATA.NEW <- GPA_UWF_ADD_DIFFGPAV1 %>% 
    filter(Fall1GPA == "Below2.77")

qplot(GPA.DATA.NEW$DIFFGPA, main = "GPA Improvement" ,fill="DIFFGPA" ) 
ggplot(GPA.DATA.NEW, aes(x=DIFFGPA)) + geom_boxplot( color="red")
x <- data.frame( GPA.DATA.NEW$ArangeSpring, GPA.DATA.NEW$BrangeSpring)
library(reshape2)
library(tidyverse)
data<- melt(x)
ggplot(data,aes(x=value, fill=variable)) + geom_density(alpha=0.25)
ggplot(data,aes(x=value, fill=variable)) + geom_histogram(alpha=0.25)
ggplot(data,aes(x=variable, y=value, fill=variable)) + geom_boxplot()
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# meanGPA1 <- GPA_UWF_ADD_DIFFGPAV0 %>% filter(Cohort <= 2018) %>%
#    summarise(meanGPA=mean(UWF1, na.rm=T))
library(stringr)
GPA_UWF_DF2 <- GPA_UWF_ADD_DIFFGPAV1 %>% 
     mutate(BEST_TEST = ifelse(BEST_TEST == 0, 0.3203125, BEST_TEST)) %>%
     mutate(is.EngineeringMajor = ifelse(str_detect(Fall1Major, "...Engineering"), "Y","N")) %>% 
     mutate(is.EngineeringMajor = ifelse(str_detect(Fall1Major, "Engineering..."), "Y",is.EngineeringMajor)) %>% 
    mutate(GPA_HIGHSCHOOL = ifelse(GPA_HIGHSCHOOL <= 3.44 , ">3.44",
                                   ifelse(GPA_HIGHSCHOOL < 3.88, "(3.44,3.88)", "3.88+"))) %>%
    mutate(HS_GPA = ifelse(GPA_HIGHSCHOOL >= 3.88 , ">=3.88", "<3.88")) %>% 
                                   #ifelse(GPA_HIGHSCHOOL < 3.88, "(3.44,3.88)", "3.88+"))) %>%
   # mutate(GPA_HIGHSCHOOL = factor(GPA_HIGHSCHOOL, levels=c( "3.44-", "(3.44,3.88)","3.88+"))) %>%
    mutate( PassedSummer = ifelse((SummerTaken == "No" & is.na(PassedSummer)), "NotAttempt","Attempt")) %>% mutate(PassedSummer =factor(PassedSummer)) %>%
    mutate(BEST_TEST = ifelse(BEST_TEST < 0.50 , "Below50%", "Above50%")) %>%
    mutate(Prior_Hours = ifelse(Prior_Hours <= 0.00,  "None",
                                ifelse(Prior_Hours < 6, "[1,6)","6+"))) %>%
    mutate(Prior_Hours = factor(Prior_Hours, levels=c( "None", "[1,6)","6+"))) %>%
    mutate(Prior_Hours = relevel(Prior_Hours, ref="6+")) %>%
     mutate(COUNTY_GROUP = ifelse(COUNTY_GROUP == "Tri-County", "Tri-County", "NonTri-County")) %>%
    # mutate(AREA_OF_STRATEGIC_EMPHASIS = ifelse(AREA_OF_STRATEGIC_EMPHASIS == "STEM", "STEM",
    #                                     ifelse(AREA_OF_STRATEGIC_EMPHASIS =="Not Available",  "Not Available", "NonSTEM"))) %>%
     mutate(STATE = ifelse( STATE == "Florida", "FL", "nonFL")) %>%
    mutate(NATION_GROUP == ifelse(NATION_GROUP == "USA", "USA", "NonUSA")) %>%
    filter(UWF1 < 2.77) %>% filter(!is.na(ArangeSpring)) %>%
    filter(!is.na(GPA_HIGHSCHOOL)) %>%
    filter(!is.na(UWF1)) %>%
    mutate(StayedSpring=ifelse(is.na(UWF2), "No","Yes")) %>%
    filter(StayedSpring != "No") # removed spring stop outs
 
MEANGPA <- GPA_UWF_DF2 %>% 
    group_by(PassedSummer) %>% dplyr::summarise(meanGPAY1 = mean(FinalY1GPA), Count=n())
MEANGPA

GPA.DATA.DIFFGPA <- GPA_UWF_DF2  %>% filter(Cohort <= 2020) %>%
    filter(!(abs(DIFFGPA - median(DIFFGPA)) > 3*sd(DIFFGPA))) %>% 
    mutate(SpringMathName = ifelse(is.na(MathnameSpring), "NotAttempt", MathnameSpring)) %>% 
    #mutate(SpringMathName = relevel(SpringMathName, ref = "NotAttempt")) %>% 
    #mutate(GPA_HIGHSCHOOL = factor(GPA_HIGHSCHOOL, levels = c("3.44-","(3.44,3.88)","3.88+"))) %>% 
    #mutate(GPA_HIGHSCHOOL = relevel(GPA_HIGHSCHOOL, ref="3.88+")) %>%
    mutate(AttemptMoreHrs=ifelse((Term2-Term1 > 0), "Yes","No")) %>% 
    select(dffGPA_varis,"Fall1AttemptHrs"=Term1, AttemptMoreHrs,"SummerTerm"=PassedSummer,  SpringMathName,is.EngineeringMajor,
          ArangeSpring, BrangeSpring,CrangeSpring,DFUSpring, HS_GPA, -GPA_HIGHSCHOOL, FinalY1GPA) %>% 
    replace(is.na(.), 0) %>%  #2189/2176/2491
    mutate(ArangeSpring = ifelse(ArangeSpring > 6, "Earned6+Hrs","Earned<6Hrs")) %>% 
    mutate(BrangeSpring = ifelse(BrangeSpring > 6, "Earned6+Hrs","Earned<6Hrs")) %>% 
    mutate(CrangeSpring = ifelse(CrangeSpring > 6, "Earned6+Hrs","Earned<6Hrs"))  
     # mutate(CrangeSpring = cut(ArangeSpring, breaks =c( -1,0, 2.99, 5.99,8.99, 30), labels=c("None",  "CRD3-","CRD[3,6)", "CRD[6,9)", "CRD9+"))) %>% 
    # mutate(BrangeSpring = cut(BrangeSpring, breaks =c( -1,0, 2.99, 5.99,8.99, 30), labels=c( "None","CRD3-","CRD[3,6)", "CRD[6,9)", "CRD9+"))) %>% 
    # mutate(CrangeSpring = cut(CrangeSpring, breaks =c( -1, 0,2.99, 5.99,8.99, 30), labels=c( "None","CRD3-","CRD[3,6)", "CRD[6,9)", "CRD9+"))) %>% 
    # mutate(DFUSpring = cut(DFUSpring, breaks =c( -1,0, 2.99, 5.99,8.99, 30), labels=c("None", "CRD3-","CRD[3,6)", "CRD[6,9)", "CRD9+")))  
    
#write.csv(GPA.DATA.DIFFGPA, "GPA.DATA.DIFFGPA.csv",row.names = F)
#addmargins(table(GPA.DATA.DIFFGPA$AttemptMoreHrs))
library(readr)
GPA.DATA.DIFFGPA <- read_csv( "GPA.DATA.DIFFGPA.csv")

Hmisc::describe(GPA.DATA.DIFFGPA)
```



```{r  plot}

mean(GPA.DATA.DIFFGPA$FinalY1GPA, na.rm=T) #0.1373989
GPA.DATA.DIFFGPAA <- GPA.DATA.DIFFGPA %>% 
    mutate(Prior_Hours = factor(Prior_Hours, levels = c("None", "[1,6)","6+")))  
library(ggpubr)
 
anovabox <- ggboxplot(GPA.DATA.DIFFGPA, x = "HS_GPA", y = "FinalY1GPA", color = "HS_GPA", 
          add = "jitter", legend = "none") +
  rotate_x_text(angle = 45)+
  geom_hline(yintercept = mean(GPA.DATA.DIFFGPA$FinalY1GPA), linetype = 2.069311)+ # Add horizontal line at base mean
  stat_compare_means(method = "anova", label.y = 3.88) +        # Add global annova p-value
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")  + xlab("High sChool GPA") + ylab("1st Year GPA")
anovabox
```


```{r  linear regression and model}
### linear equation
library(ggpmisc)
con_p <- GPA_UWF_MAAV_majorAPR %>% filter(GPA_HIGHSCHOOL != 0.00) %>% 
    filter(UWF1 != 0.00) %>% mutate(logUWF1=log(UWF1)) %>% 
     filter(!(abs(GPA_HIGHSCHOOL - median(GPA_HIGHSCHOOL)) > 3*sd(GPA_HIGHSCHOOL)))  
formula <- y ~ x
mlplot <- ggplot(con_p, aes(x= UWF1, y= UWF2, color = GPA_HIGHSCHOOL)) +
  geom_point(alpha = 0.3 ) +
  #facet_wrap(~clarity, scales = "free_y") +
  geom_smooth(method = "lm", formula = formula, se = F) +
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               label.x.npc = "right", label.y.npc = 0.15,
               formula = formula, parse = TRUE, size = 5, color="red") +
    labs(title = "Fall GPA vs. Spring GPA") +xlab("Fall GPA")+ylab("Spring GPA")
mlplot


 
```





## Methods

-Focus on regression models to predict continuous dependent variable.
1. Multiple linear regression
2. Regularized regression models
    
**Multiple Linear Regression **

Prefer using a multiple linear regression model since it is more realistic -things often depend on 2, 5, 10 or even more factors.

-Evaluate performance of the model considering R-squared values and Root Mean Squared Error (RMSE)
-Higher R-squared score and lower RMSE values indicate better model.
-The R-squared measures how much of the total variability is explained by our model
-Multiple regressions are always better than simple ones, as with each additional variable you add, the explanatory power may only increase or stay the same
-adj.R^2 < R^2
-adj.R^2 is a measure that how well your model fits the data. However, it penalizes the use of variables that are meaningless for the regression.
-Add insignificant variables lose explanatory power. adj.R^2 basis for comparing models.(same variables and data set)
-IF adding a new parameter increases R-squared but decreases adj. R-squared, the variable can be omitted since it holds no predictive power.
-F-statistics: used for testing the overall significant of the model
-F-test(p-value): Ho-all betas are zero
-How to asses
-Assumptions: linearity (linear regression), there are more sophisticated models, No endogeneity(situations in which an explanatory variable is correlated with the error term), normality and homoscedastictiy (errors are normaly distributed), no autocorrelation, no mulitcollinearity
    
**Ridge Regression**

Extension of linear regression where the loss function is modified to minimize the complexity of the model. 
This modification is done by adding a penalty parameter that is equivalent to the square of the magnitude of the coefficients.

 

```{r pacages, message=FALSE, warning=FALSE, include=FALSE}
library(plyr)
library(readr)
library(dplyr)
library(caret)
library(ggplot2)
library(repr)
library(corrgram)
library(corrplot)
library(mlbench)
library(psych)
library(readr)
# GPAY1_MAA2022_DF_V2 <- GPAY1_MAA2022_DF_V2 %>% 
#      filter(!is.na(GPA_ENTERING_SECOND_FALL)) %>% 
#     filter(!is.na(GPA_HIGHSCHOOL)) %>% filter(!is.na(total_points_earned)) 
# #%>%  #filter(AveGPA2.77 =="No") 
# #colSums(is.na(GPAY1_MAA2022_DF_V2))
colnames(GPA_UWF_ADD_DIFFGPAV1)
```

### Data Partitioning/Scaling/Correlation Matrix

```{r}
# num.cols <- GPA.DATA %>% 
#     select(where(is.numeric)) %>% 
#     select(-FinalY1GPA,-COHORT_YEAR )# collinearity

num.cols.raw <- GPA_UWF_ADD_DIFFGPAV1 %>% select(6,7,8,10:15,22,24,26,32,42,48,51,53:57) %>% replace(is.na(.),0) %>% 
    select(where(is.numeric)) 
# colSums(is.na(num.cols))
# GPA.DATA[,num.col.names] <- predict(prep_num.cols, GPA.DATA[,num.col.names])
# summary(GPAY1_MAA2022_DF_V2)

#num.cols.cor <- num.cols %>% na.omit()
nu.raw.cor <- num.cols.raw  %>%  na.omit()
# colSums(is.na(num.cols.cor))
#num.cols.corr <- cor(num.cols.cor)
num.cols.corr.raw <- cor(nu.raw.cor)

#num.cols.corr
#corrplot(num.cols.corr, method = "pie" ) # too small
#psych::pairs.panels(num.cols.corr, cex.cor=2, main="Histograms and Correlations for a Data Matrix")
psych::pairs.panels(num.cols.corr.raw, cex.cor=4, main="Histograms and Correlations for a Data Matrix")

#qplot(GPA.DATA$FinalY1GPA, main = "1st-Year GPA for At-risk FTIC")

```
### Data Partition

```{r}
# colSums(is.na(GPAY1_MAA2022_DF_V2)) 
set.seed(123)
id <- sample(2, nrow(GPA.DATA.DIFFGPA), replace = T, prob = c(0.7,0.3))
train_t <- GPA.DATA.DIFFGPA[id ==1, ]
test_t <- GPA.DATA.DIFFGPA[id ==2, ]

# Custom control parameters
custom1 <- trainControl(method = "repeatedcv",number = 10, repeats = 3, verboseIter = T)


```





### Regression

```{r message=FALSE, warning=FALSE, include=FALSE}
#colSums(is.na(GPAY1_MAA2022_DF_V2))
lm.1 <- train(FinalY1GPA ~ ., data = train_t, method ="lm", trControl = custom1)
lm.1$results #RMSE 0.5282094 /0.7067871
 
summary(lm.1)
summary(m1 <- lm( FinalY1GPA ~., data= train_t))
car::vif(m1)
summary(m2 <- step(m1, direction = "backward"))
sqrt(mean(m2$residuals^2))  #RMSE0.5191501 0.7238,

#Ridge
ridge.1 <- train(FinalY1GPA ~ ., data = train_t, method ="glmnet", 
                 tuneGrid = expand.grid(alpha = 0, lambda = seq(0.0001, 0.5,length = 5)), # increase lambda increase fananltyhrink 
                 trControl = custom1) # best  lambda = 1e-04 0.0004

ridge.1
# Lasso
set.seed(213)
lasso.1 <- train(FinalY1GPA ~ ., data = train_t, method ="glmnet", 
                 tuneGrid = expand.grid(alpha = 1, lambda = seq(0.0001, 0.5,length = 5)), # increase lambda increase fananltyhrink 
                 trControl = custom1)

#Elastic Net
elasticnet.1 <- train(FinalY1GPA ~ ., data = train_t, method ="glmnet", 
                 tuneGrid = expand.grid(alpha = seq(0, 1, length = 5), lambda = seq(0.0001, 0.5,length = 10)),  trControl = custom1)
#alpha = 0.111, lambda = 1e-04 
```

### Plots

```{r}
library(vip)
p1 <- vip(m1, num_features = length(coef(m1)), geom = "point", horizontal = FALSE)
p2 <- vip(m1, num_features = length(coef(m1)), geom = "point", horizontal = TRUE, mapping = aes_string(color="Sign"))
grid.arrange(p1, p2, nrow=1)
factors <-  m1$coefficients   %>%  data.frame() 
factors$names <-  rownames(factors)
factors$Coef <-  round(factors$. ,8)
factros <- factors %>% arrange(Coef) %>% filter(Coef>0)

p_backward <- ggplot(factors[-1,], aes(y =  factor(names), x = Coef , colour=factor(sign(Coef)), group=names, label = names)) + 
    geom_point(size=2.5) + geom_text(size=2.5, vjust = 0, hjust = 0) + theme(legend.position = "none")
p_backward  
plot(lm.1$finalModel)

plot(ridge.1) #higher lambda increase errors
plot(ridge.1$finalModel, xvar= "lambda") #relax lambda coefficient smaller
plot(ridge.1$finalModel, xvar ="dev", label = F)
plot(varImp(ridge.1, scale = F), main="Variable Importance using Ridge Regression")
plot(lasso.1)
plot(lasso.1$finalModel, xvar= "lambda") #relax lambda coefficient smaller
plot(lasso.1$finalModel, xvar ="dev", label = F) # show overfitting at the end
plot(varImp(lasso.1, scale = F),main="Variable Importance using Lasso Regression")
plot(elasticnet.1)
plot(elasticnet.1$finalModel, xvar= "lambda") #relax lambda coefficient smaller
plot(elasticnet.1$finalModel, xvar ="dev", label = F) # show overfitting at the end
plot(varImp(elasticnet.1, scale = F), main="Variable Importance using Elastic Net Regression")
p2
```





```{r}

#comparison
all_models <- list(LinerModel = lm.1, RidgeModel = ridge.1, LassoModel = lasso.1, ElasticNetModel =elasticnet.1)
res <- resamples(all_models)
summary(res)
bwplot(res)
xyplot(res, metric = "RMSE")

#best model
# lasso.1$bestTune
# lasso.1$bestTune
best.M <- lasso.1$finalModel
coef(best.M, s = lasso.1$bestTune$lambda)
saveRDS(lasso.1,"FinalModel.rds")
final.model <- readRDS("FinalModel.rds")

#prediction
 (summary(res))
Pred.train <- predict(final.model, train_t)
sqrt(mean((train_t$FinalY1GPA - Pred.train)^2)) # 0.415200
Pred.test <- predict(final.model, test_t)
sqrt(mean((test_t$FinalY1GPA - Pred.test)^2)) # 0.4003097

```



 
